Over Twenty Years of Virtual Machine Development and Debugging Through Simulation

### 2018-08-14 paper.tex
fix;

HERE
\paragraph{Simulation.} Once we had partially written the new Sweep algorithm, we started the simulator. Since the algorithm was partially written, we had to write the missing pieces inside the debugger, installing the new code at runtime, as one can do it in Smalltalk. In addition, the simulator has an interesting property: each time it performs a GC, either a scavenge or an old space collection, it first copies the simulated memory (\emph{i.e.,} the heap), performs the GC there, and if no assertion fails, it then performs the GC on the original version. This means that if the algorithm was not working, an assertion would fail or an error such as incorrect memory access would happen in a copy of the heap. The simulator can then reproduce the exact same crash as many times as we want duplicating again and again the same original heap. This is very convenient to debug specific GC bugs because often in GC bugs, when the bug happens, the memory is already corrupted and it's quite difficult to track back where the bug comes from without rewinding the memory state.

@@

A few years ago, we added support in the VM for read-only objects \cite{ROObject}. Read-only objects were critical performance-wise for specific customers using them in the context of object databases. To maximize the performance, we changed the templates in the JIT compiler for the different memory stores. Since the template JIT has to generate very quickly machine code, it goes through the code three times: (1) it scans the bytecode, (2) it generates abstract instructions from the bytecode and (3) it generates the native instructions from the abstract instructions. The abstract instructions are mapped almost one to one to machine instructions, their main purpose is to compact the generated code by removing Nops used for jump targets or finding out out the size of jumps.

>>>>

\paragraph{Simulation.} Once we had partially written the new Sweep algorithm, we started the simulator. Since the algorithm was partially written, we had to write the missing pieces inside the debugger, installing the new code at runtime, as one can do it in any Smalltalk application. In addition, the simulator has an interesting property: each time it performs a GC, either a scavenge or an old space collection, it first copies the simulated memory (\emph{i.e.,} the heap), performs the GC there, and if no assertion fails, it then performs the GC on the original version. This means that if the algorithm was not working, and likely an assertion would fail or an error such as incorrect memory access would be raised, the erratic behavior would have happened in a copy of the heap. The simulator can then reproduce the exact same erratic behavior as many times as the VM developer wants duplicating again and again the same original heap. This is very convenient to debug specific GC bugs. Indeed, often in GC bugs, when the bug happens, the memory is already corrupted and it's quite difficult to track back where the bug comes from without rewinding the memory state, which is usually very difficult or impossible.

@@

HERE
A few years ago, we added support in the VM for read-only objects \cite{ROObject}. Read-only objects were critical performance-wise for specific customers using them in the context of object databases. To maximize the performance, we changed the templates in the JIT compiler for the different memory stores. Since the template JIT has to generate very quickly machine code, it goes through the code three times: (1) it scans the bytecode, (2) it generates abstract instructions from the bytecode and (3) it generates the native instructions from the abstract instructions. The abstract instructions are mapped almost one to one to machine instructions, their main purpose is to compact the generated code by removing Nops used for jump targets or finding out out the size of jumps.
